Acceleration of Derivative Calculations with Application to Radial Basis Function -- Finite-Differences on the Intel MIC Architecture

### 2014-04-16 acm/ieee_paper_final.tex
Change section prior to conclusion. The numbers to refect the fact that the cache of Cascade, E5-3670 has 512 bit cache line (64 bytes). .

\begin{figure}[t]
\centering
\includegraphics[width=0.65\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
\caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square. In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
\end{figure}

@@

\begin{figure*}[tbh]
  \centering
  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}

@@

\begin{figure*}
% Figure 9
  \centering
% was width=.48
  \subfigure[Supercompact and Random.]{\includegraphics[width=.4\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
% was width=.48
  \subfigure[Compact and RBFFD.]{\includegraphics[width=.4\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}

@@

  \includegraphics[width=.9\linewidth]{figures/mic_performance_nb_threads.pdf}

@@

\ger{The main idea of this paper, computing 16 derivatives at a time,
can be applied on different architectures such as multiprocessor
multicore or GPU. To illustrate this fact, we conducted similar experiments
on Cascade at the Minnesota Supercomputing Institute
using two Intel Xeon E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
cache per core, with hyperthreading disabled.) The cores support the AVX instruction set. 
Both the compiler-generated code and the code written with the 
help of the AVX instruction set reach a read bandwidth of
85~GB/s. Using cache lines of 256 bits, we find a best/worst speeds of 4/20\; Gflops/s and 
50/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 41\% of peak theoretical 
performance on this particular processor setup. For references, we achieve 
140/210=66\% of peak theoretical performance on the MIC. Note that peak performance 
is not that of the processor, but of the best possible algorithm on the processor. 
The gain in the 4/4 case with respect to 4x 1/1 case is respectively 4x on the MIC 
and 2.5x on the front end.  
}

@@

Our implementation makes use of the IMCI MIC instruction set and
includes a number of swizzling and channel swapping operations, for an
extremely efficient tensor product implementation. 
\NOTE{From Evan: this is the first time $96^3$ is mentionned. Might wish to mention 
it earlier.} On a $96^3$ 3D
grid, it reaches a performance of 140~Gflop/s which is 2.7 times
faster than the best possible performance achievable for a single
vector and a single derivative and more than 7 times faster than a
practical implementation.

>>>>

\begin{figure}[t]
\centering
\includegraphics[width=0.60\linewidth]{figures/RBFStencil_n32_a}\label{fig:stencil}
\caption{An example of an RBF-FD stencil of size $n=32$ on a sphere of $N=1024$ nodes to approximate any derivative operator at the point in the square. In nonsparse format, the DM  is $1024\times1024$ with 32 nonzero entries per row.}
\end{figure}

@@

\begin{figure*}[tbh]
  \centering
  \subfigure[Best case.]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
  \subfigure[Worst case.]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.28\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}

@@

\begin{figure*}
% Figure 9
  \centering
% was width=.48
  \subfigure[Supercompact and Random.]{\includegraphics[width=.35\linewidth]{figures/random_supercompact.pdf}\label{fig:perf_ideal}}
% was width=.48
  \subfigure[Compact and RBFFD.]{\includegraphics[width=.35\linewidth]{figures/rbf_compact.pdf}\label{fig:perf_realistic}}

@@

  \includegraphics[width=.8\linewidth]{figures/mic_performance_nb_threads.pdf}

@@

\ger{The main idea of this paper, computing 16 derivatives at a time,
can be applied on different architectures such as multiprocessor
multicore or GPU. To illustrate this fact, we conducted similar experiments
on Cascade at the Minnesota Supercomputing Institute
using two Intel Xeon E5 2670 nodes (8 cores, 2 threads and 20 Mbytes of L2
cache per core, with hyperthreading disabled.) The cores support the AVX instruction set. 
Both the compiler-generated code and the code written with the 
help of the AVX instruction set reach a read bandwidth of
85~GB/s. Using cache lines of 512 bits, we find a best/worst speeds of 2.4/20\; Gflops/s and 
32/120\; Gflops/s, respectively for the 1vec/1mat and 4vec/4mat cases. 
Our implementation on Cascade reaches 45 Gflops for the 4/4\, case, or 45/120=37\% of peak theoretical 
performance on this particular processor setup. For references, we achieve 
140/210=66\% of peak theoretical performance on the MIC. Note that peak performance 
is not that of the processor, but of the best possible algorithm on the processor. 
The gain in the 4/4 case with respect to 4x 1/1 case is respectively 4x on the MIC 
and 2.5x on the front end.  
}

% cascade
%[  2.35701906   8.04733728   9.37931034  31.62790698  36.75675676]
%[  20.60606061   32.96969697   75.55555556  120.88888889  226.66666667]

@@

Our implementation makes use of the IMCI MIC instruction set and
includes a number of swizzling and channel swapping operations, for an
extremely efficient tensor product implementation. 
\NOTE{From Evan: this is the first time $96^3$ is mentionned. Might wish to mention 
it earlier.} On a $96^3$ 3D
grid, it reaches a performance of 140~Gflop/s which is 2.7 times
faster than the best possible performance achievable for a single
vector and a single derivative and more than 7 times faster than a
best case practical 1/1 implementation.
